\documentclass[../main.tex]{subfiles}

\begin{document}
\doublespacing
\singlespacing
\chapter*{Acknowledgements}
\doublespacing

I would like to thank my supervisors, Dr. Glenn Harrison and Dr. Don Ross, for their tremendous support and guidance throughout this process.
I know I would not have wanted to do this thesis with any other supervisors.
Their tutelage has given me an invaluable first step into my career as an economist.
Also, a strangely debilitating fear of red ink and cursive.

I came to find that writing a thesis focused on such a specific topic as stochastic models can be an intellectually lonely endeavour.
I'd like to thank my good friend and economist Dr. Andre Hofmeyr for his friendship throughout this process and always being available to talk economics over beer and hikes.
His expertise helped refine some of my better ideas and weed out the weaker ones.

To Dr. Lisa Rutstr\"{o}m, for recruiting me to run her experiments at the University of Central Florida, convincing me to continue my academic career by pursuing my masters at Georgia State University and finally encouraging me to pursue a PhD at the University of Cape Town.
Indeed, this thesis would never have been written had she not encouraged me to do more and be better at every stage of my academic career.
Thank you.

My wife, Quanita, for helping me though the joy and heartbreak of the aforementioned red pen and cursive comments and having the grace to smile and nod when I used her as a sounding board.

\end{document}
